\huge \textbf{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}

### 2025-04-29 paper/paper_draft.tex
paper changes

% \section{Fisher matrices raw data}

% Bispectra Fisher matrices:\\
% \begin{gather*}
% F^B_{\psi_c} =
% \begin{pmatrix}
% 4\times 10^{-3} & 6\times 10^{-4} & -6\times 10^{-3} & -9\times 10^{-3} & 1\times 10^{-4} & -4\times 10^{-3} & -1\times 10^{-3}\\
% 6\times 10^{-4} & 1\times 10^{-4} & -1\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-5} & -6\times 10^{-4} & -2\times 10^{-4}\\
% -6\times 10^{-3} & -1\times 10^{-3} & 1\times 10^{-2} & 1\times 10^{-2} & -2\times 10^{-4} & 6\times 10^{-3} & 2\times 10^{-3}\\
% -9\times 10^{-3} & -2\times 10^{-3} & 1\times 10^{-2} & 2\times 10^{-2} & -2\times 10^{-4} & 8\times 10^{-3} & 3\times 10^{-3}\\
% 1\times 10^{-4} & 2\times 10^{-5} & -2\times 10^{-4} & -2\times 10^{-4} & 3\times 10^{-6} & -1\times 10^{-4} & -3\times 10^{-5}\\
% -4\times 10^{-3} & -6\times 10^{-4} & 6\times 10^{-3} & 8\times 10^{-3} & -1\times 10^{-4} & 4\times 10^{-3} & 1\times 10^{-3}\\
% -1\times 10^{-3} & -2\times 10^{-4} & 2\times 10^{-3} & 3\times 10^{-3} & -3\times 10^{-5} & 1\times 10^{-3} & 4\times 10^{-4}
% \end{pmatrix},
% \\
% F^B_{\psi_s} =
% \begin{pmatrix}
% 3\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -1\times 10^{-1} & 5\times 10^{-3} & -1\times 10^{-1} & -1\times 10^{-1}\\
% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 5\times 10^{-4} & -1\times 10^{-2} & -1\times 10^{-2}\\
% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
% -1\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
% 5\times 10^{-3} & 5\times 10^{-4} & -3\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 6\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 5\times 10^{-2} & -2\times 10^{-3} & 5\times 10^{-2} & 5\times 10^{-2}
% \end{pmatrix},
% \\
% F^B_{\psi_{s + c}} =
% \begin{pmatrix}
% 4\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -2\times 10^{-1} & 5\times 10^{-3} & -2\times 10^{-1} & -1\times 10^{-1}\\
% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 6\times 10^{-4} & -2\times 10^{-2} & -1\times 10^{-2}\\
% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 2\times 10^{-1} & -4\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
% -2\times 10^{-1} & -2\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 8\times 10^{-2} & 7\times 10^{-2}\\
% 5\times 10^{-3} & 6\times 10^{-4} & -4\times 10^{-3} & -3\times 10^{-3} & 7\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
% -2\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 8\times 10^{-2} & -2\times 10^{-3} & 7\times 10^{-2} & 6\times 10^{-2}\\
% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}
% \end{pmatrix}.
% \end{gather*}

% Powerspectra Fisher matrices:
% \begin{gather*}
% F^P_{\psi_c} =
% \begin{pmatrix}
% 2\times 10^{-2} & 4\times 10^{-3} & -4\times 10^{-2} & 8\times 10^{-3} & 9\times 10^{-4} & -3\times 10^{-2} & -1\times 10^{-2}\\
% 4\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-4} & -5\times 10^{-3} & -2\times 10^{-3}\\
% -4\times 10^{-2} & -9\times 10^{-3} & 8\times 10^{-2} & -9\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-2} & 2\times 10^{-2}\\
% 8\times 10^{-3} & -6\times 10^{-4} & -9\times 10^{-3} & 2\times 10^{-2} & 3\times 10^{-4} & -1\times 10^{-2} & -4\times 10^{-3}\\
% 9\times 10^{-4} & 2\times 10^{-4} & -2\times 10^{-3} & 3\times 10^{-4} & 3\times 10^{-5} & -1\times 10^{-3} & -4\times 10^{-4}\\
% -3\times 10^{-2} & -5\times 10^{-3} & 6\times 10^{-2} & -1\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 2\times 10^{-2}\\
% -1\times 10^{-2} & -2\times 10^{-3} & 2\times 10^{-2} & -4\times 10^{-3} & -4\times 10^{-4} & 2\times 10^{-2} & 6\times 10^{-3}
% \end{pmatrix},
% \\
% F^P_{\psi_s} =
% \begin{pmatrix}
% 2\times 10^{-1} & 1\times 10^{-2} & -1\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-3} & -6\times 10^{-2} & -7\times 10^{-2}\\
% 1\times 10^{-2} & 1\times 10^{-3} & -9\times 10^{-3} & -3\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -6\times 10^{-3}\\
% -1\times 10^{-1} & -9\times 10^{-3} & 7\times 10^{-2} & 2\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 4\times 10^{-2}\\
% -3\times 10^{-2} & -3\times 10^{-3} & 2\times 10^{-2} & 9\times 10^{-3} & -4\times 10^{-4} & 1\times 10^{-2} & 1\times 10^{-2}\\
% 2\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -4\times 10^{-4} & 3\times 10^{-5} & -7\times 10^{-4} & -8\times 10^{-4}\\
% -6\times 10^{-2} & -5\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -7\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}\\
% -7\times 10^{-2} & -6\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -8\times 10^{-4} & 2\times 10^{-2} & 3\times 10^{-2}
% \end{pmatrix},
% \\
% F^P_{\psi_{s+c}} =
% \begin{pmatrix}
% 9\times 10^{-2} & 9\times 10^{-3} & -7\times 10^{-2} & -1\times 10^{-2} & 1\times 10^{-3} & -4\times 10^{-2} & -4\times 10^{-2}\\
% 9\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -4\times 10^{-3}\\
% -7\times 10^{-2} & -9\times 10^{-3} & 7\times 10^{-2} & 8\times 10^{-3} & -1\times 10^{-3} & 5\times 10^{-2} & 3\times 10^{-2}\\
% -1\times 10^{-2} & -2\times 10^{-3} & 8\times 10^{-3} & 2\times 10^{-2} & -7\times 10^{-5} & -3\times 10^{-4} & 5\times 10^{-3}\\
% 1\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -7\times 10^{-5} & 3\times 10^{-5} & -9\times 10^{-4} & -6\times 10^{-4}\\
% -4\times 10^{-2} & -5\times 10^{-3} & 5\times 10^{-2} & -3\times 10^{-4} & -9\times 10^{-4} & 3\times 10^{-2} & 2\times 10^{-2}\\
% -4\times 10^{-2} & -4\times 10^{-3} & 3\times 10^{-2} & 5\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}
% \end{pmatrix}.
% \end{gather*}

>>>>

\section{Numerical derivative}
The derivatives with respect to cosmological parameters were taken with a central difference formula, i.e.
\begin{equation*}
    f'(x) = \frac{f(x+h) - f(x-h)}{2h} + \mathcal O (h^2).
\end{equation*}
Each change of the cosmological parameters requires a recalculation of the entire cosmology, making it computationally expensive. For the approximation to be accurate a balance needs to be found between numerical errors for small $h$ and a larger $\mathcal O(h^2)$ error for larger $h$. The $h$ values chosen for each parameter are shown in table \ref{tab:cosmo-params-diff} and are similar to the values used in \cite{Namikawa_2016}\footnote{As confirmed during a conversation with the author.}.

\begin{table}[h]
    \centering
    \begin{tabular}{lcc}
    \toprule
    Parameter & Fiducial value & Finite difference ($h$) \\
    \midrule
    $H$            & $67.4$               & fiducial\,$\times0.1$   \\
    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.1$   \\
    $\Omega_c h^2$ & $0.119$              & fiducial\,$\times0.005$ \\
    $n_s$          & $0.965$              & fiducial\,$\times0.005$ \\
    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.1$   \\
    $m_\nu$        & $0.06$               & fiducial\,$\times0.1$   \\
    $w_0$          & $-1$                 & $0.03$                  \\
    \bottomrule
    \end{tabular}
    \caption{Fiducial cosmological parameters and their finite-difference steps}
    \label{tab:cosmo-params-diff}
\end{table}

To test the accuracy, we vary $h$ by $\pm 5\%$ and $\pm 10\%$ and plot the relative change in the derivative of the equilateral bispectrum in figure \ref{fig:deraccuracy}. As long as numerical noise doesn't dominate, it can be shown that the relative error in our approximation is approximately 5 times the relative difference that taking $h \rightarrow h(1 \pm 0.1)$ leads to. We thus conclude that, based on this test, the derivatives are almost always computed with up to one percent of error.

\begin{figure}[t]
    \centering
    \includegraphics[height=\textheight]{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/plots/derivative_accuracy.png}
    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
    \label{fig:deraccuracy}
\end{figure}
